---
title: "Ggplot_part3"
author: "Zhang Xiaoqiong"
date: "1/22/2020"
output:
  pdf_document: default
  html_document: default
---
## Ggplot Part3 
Plots for different use.

```{r dataset}
options(scipen = 999)
library(ggplot2)
theme_set(theme_bw())
data("midwest",package = "ggplot2")
data("mpg",package = "ggplot2")
data("economics",package = "ggplot2")
dim(midwest)
dim(mpg)
```
### 1. Correlation
The following plots help to examine how well correlated two variables are.

#### Scatterplot (midwest)
It van be drawn using geom_point(); geom_smooth draws a smoothing line (based on loess) by setting method="lm"
```{r}
gg = ggplot(midwest,aes(x=area,y=poptotal)) +
  geom_point(aes(col=state,size=popdensity)) +
  geom_smooth(method="loess",se=F) +
  xlim(c(0,0.1)) +
  ylim(c(0,500000)) +
  labs(subtitle = "Area Vs Population",
       y= "Poplation",
       x = "Area",
       title = "Scatterplot",
       caption = "Source: midwest")

plot(gg)
```

#### Scatterplot With Encircling

Encircling certain special group of points or region can be done using the geom_encircle() in *ggalt* package
```{r}
library(ggalt)
# Encircling Area Data
midwest_select = midwest[midwest$poptotal > 350000 &
                           midwest$poptotal <= 500000 &
                           midwest$area > 0.01 &
                           midwest$area < 0.1, ]
# plot
gg + geom_encircle(aes(x=area,y=poptotal),
                   data = midwest_select,
                   color = "red",
                   size = 2,
                   expand = 0.08) +
  labs(title = "Scatterplot + Encircle")
```

#### Jitter Plot (mpg example)
Method 1 to deal with overlapping points
```{r}
g = ggplot(mpg,aes(cty,hwy))

# Scatter plot
g + geom_point() + geom_smooth(method="lm",se=F) +
  labs(subtitle = "mpg: city vs highway mileage",
       y = "hwy",
       x = "cty",
       title = "Scatterplot with overlapping points",
       caption = "Source: midwest")

# Jitter plot
g + geom_jitter(width=.5,size=1) + 
  labs(title = "Jittered points",
       subtitle = "mpg: city vs highway mileage",
       y = "hwy",
       x = "cty")
```
#### Counts Chart (mpg example)
Second method to overcome the problem of data points overlap
```{r}

# Jitter plot
g + geom_count(col="tomato3",show.legend = F) + 
  labs(title = "Counts Plot",
       subtitle = "mpg: city vs highway mileage",
       y = "hwy",
       x = "cty")
```
#### Bubble Charts (mpg example)
Bubble charts provide a better visual comparison between groups based on: a categorical variable (by changing the color) and another continuous variable (by changing the size of points)
```{r}

# dataset 
mpg_select = mpg[mpg$manufacturer %in% c("audi","ford","honda","hyundai"),]

g = ggplot(mpg_select,aes(displ,cty)) +
  labs(subtitle = "mpg: Displacement vs City Mileage",
       title = "Bubble chart")

g + geom_jitter(aes(col=manufacturer, size=hwy)) +
  geom_smooth(aes(col=manufacturer),method="lm",se=F)
```
#### Boxplot / Marginal Histogram (mpg example)
It shows the relationship as well as the distributin in the same chart.
It can be done with *ggMarginal()* function from the *ggExtra* package
```{r}
library(ggExtra)

g = ggplot(mpg,aes(cty,hwy)) + geom_count() + geom_smooth(method="lm",se=F)

ggMarginal(g,type="histogram",fill="transparent")
ggMarginal(g,type="boxplot",fill="transparent")

```

#### Correlogram (mecars example)
It examines the corellation of ultiplecontinuous variables present in the same dataframe
```{r}
library(ggcorrplot)

#data
data("mtcars")
corr = round(cor(mtcars),1)

# plot
ggcorrplot(corr,hc.order=TRUE,
           type = "lower",
           lab = TRUE,
           lab_size = 3,
           method = "circle",
           colors = c("tomato2","white","springgreen3"),
           title="Correlogram of mtcars",
           ggtheme = theme_bw)
```

### 2.Deviation 

#### Diverging bars (mtcars example)
It can handle both negative and positive values with function *geom_bar()*.
Note: by default, geom_bar() has the *stat* set to count, which is to make histogram. To make bar charts, *stat* needs to be set to *identity*

```{r}
library(ggcorrplot)

# data prep
mtcars$Car_name = rownames(mtcars) # create new column for car names
mtcars$mpg_z = round((mtcars$mpg - mean(mtcars$mpg))/sd(mtcars$mpg),2) # normalize the data
mtcars$mpg_type = ifelse(mtcars$mpg_z < 0,"below","above")
mtcars = mtcars[order(mtcars$mpg_z), ] # sort
mtcars$Car_name = factor(mtcars$Car_name,levels=mtcars$'car name') # convert to factor to retain sorted order in plot

# Diverging Barcharts
ggplot(mtcars,aes(x=Car_name,y=mpg_z,label=mpg_z)) +
  geom_bar(stat='identity',aes(fill=mpg_type),width=.5) +
  scale_fill_manual(name="Mileage",
                    labels = c("Above Average","Below Average"),
                    values = c("above" = "#00ba38","below" = "#f8766d")) +
  labs(subtitle = "Normalised mileage from 'mtcars'",
       title = "Diverging Bars") +
  coord_flip()
```

#### Diverging Lollipop Chart
It can handle both negative and positive values with function *geom_bar()*.
Note: by default, geom_bar() has the *stat* set to count, which is to make histogram. To make bar charts, *stat* needs to be set to *identity*

```{r}
ggplot(mtcars,aes(x=Car_name,y=mpg_z,label=mpg_z)) +
  geom_point(stat='identity',fill="black",size=6) +
  geom_segment(aes(y = 0,
                   x = Car_name,
                   yend = mpg_z,
                   xend = Car_name),
               color = "black") +
  geom_text(color = "white", size = 2) +
  labs(title="Diverging Lollipop Chart",
       subtitle = "Normalized mileage from 'mtcars': Lollipop") +
  ylim(-2.5,2.5) +
  coord_flip()
```

#### Diverging Dot Plot

```{r}
ggplot(mtcars,aes(x=Car_name,y=mpg_z,label=mpg_z)) +
  geom_point(stat='identity',aes(col=mpg_type),size=6) +
  scale_color_manual(name="Mileage",
                     labels = c("Above Average","Below Average"),
                     values = c("above" = "#00ba38","below" = "#f8766d")) +
  geom_text(color = "white", size = 2) +
  labs(title="Diverging Dot Plot",
       subtitle = "Normalized mileage from 'mtcars': Dotplot") +
  ylim(-2.5,2.5) +
  coord_flip()
```

#### Area Chat (economics example)
It typically used to visualize how a particular metric (such as % return from a stock) performed compared to a certain baseline.
```{r}
library(quantmod)
library(zoo)
library(xts)
# Compute % Return
economics$returns_perc = c(0,diff(economics$psavert)/economics$psavert[-length(economics$psavert)])

# Create break points and labels for axis ticks
brks = economics$date[seq(1,length(economics$date),12)]
lbls = lubridate::year(economics$date[seq(1,length(economics$date),12)])

#plot
ggplot(economics[1:100,],aes(date,returns_perc)) +
  geom_area() +
  scale_x_date(breaks=brks,labels=lbls) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Area Chart",
       subtitle = "Perc Returns for Personal Savings",
       y = "% Returns for Personal savings",
       caption = "Source: economics")
```

### 3.Ranking 

#### Ordered Bar Chart (mpg)

```{r}
# Prepare data: grow mean sity mileage by manufacturer
cty_mpg = aggregate(mpg$cty,by=list(mpg$manufacturer),FUN=mean) #aggregate
colnames(cty_mpg) = c("make","mileage") # Change the column names
cty_mpg = cty_mpg[order(cty_mpg$mileage),] # sort
cty_mpg$make = factor(cty_mpg$make,levels = cty_mpg$make) # to retain the order in the plot
head(cty_mpg,4)
```
```{r}
# plot
library(ggplot2)
theme_set(theme_bw())

ggplot(cty_mpg,aes(x=make, y=mileage)) +
  geom_bar(stat = "identity",width = .5,fill = "tomatos3") +
  labs(title = "Ordered Bar Chart",
       subtitle = "Make Vs Avg. Mileage",
       caption = " Source: mpg") +
  theme(axis.text.x = element_text(angle=65,vjust=0.6))

```

#### Lollipop Chart (mpg)

```{r}
ggplot(cty_mpg,aes(x=make, y=mileage)) +
  geom_point(size=3) +
  geom_segment(aes(x=make,
                   xend=make,
                   y=0,
                   yend=mileage)) +
  labs(title = "Lollipop Chart",
       subtitle = "Make Vs Avg. Mileage",
       caption = " Source: mpg") +
  theme(axis.text.x = element_text(angle=65,vjust=0.6))

```

#### Dot Plot (mpg)

```{r}
library(scales)
theme_set(theme_classic())

# plot
ggplot(cty_mpg,aes(x=make, y=mileage)) +
  geom_point(size=3,col="tomato2") +
  geom_segment(aes(x=make,
                   xend=make,
                   y=min(mileage),
                   yend=max(mileage)),
               linetype = "dashed",
               size = 0.1) +
  labs(title = "Dot Plot",
       subtitle = "Make Vs Avg. Mileage",
       caption = " Source: mpg") +
  coord_flip()

```

#### Slope Chart 

```{r}
# Prep data
df = data.frame(continent=c("Africa","Americas","Asia","Europe","Oceania"),
                `1952` = c(1252.57246582115,4079.0625522,5195.48400403939,5661.05743476,10298.08565),
                `1957` = c(1385.23606225577,4616.04373316,4003.13293994242,6963.01281593333,11598.522455))

left_label = paste(df$continent,round(df$X1952),sep=",")
right_label = paste(df$continent,round(df$X1957),sep=",")
                    
df$class = ifelse((df$X1957 - df$X1952)<0,"red","green")
```

```{r}
#plot
p = ggplot(df) + 
  geom_segment(aes(x=1,xend=2, y = X1952,yend=X1957,col=class),size=.75,show.legend = F) +
  geom_vline(xintercept = 1,linetype="dashed",size=.1) +
  geom_vline(xintercept = 2,linetype="dashed",size=.1) +
  scale_color_manual(labels=c("Up","Down"),
                     values=c("green"="#00ba35","red"="#f8766d")) +
  labs(x="",y="Mean GdpPerCap") +
  xlim(.5,2.5) + 
  ylim(0, (1.1*(max(df$X1952,df$X1957))))

# Add texts
p = p + geom_text(label=left_label,y=df$X1952,x=rep(1,NROW(df)),hjust=1.1,size=3.5)
p = p + geom_text(label=right_label,y=df$X1957,x=rep(2,NROW(df)),hjust=-0.1,size=3.5)
p = p + geom_text(label="Time 1", x=1, y =1.1*(max(df$X1952,df$X1957)),hjust=1.2,size=5)
p = p + geom_text(label="Time 2", x=2, y =1.1*(max(df$X1952,df$X1957)),hjust=-0.1,size=5)
# Minify theme
p + theme(panel.background = element_blank(),
          panel.grid = element_blank(),
          axis.ticks = element_blank(),
          axis.text.x = element_blank(),
          panel.border = element_blank(),
          plot.margin = unit(c(1,2,1,2),"cm"))


```

### 4.Distribution

#### Histogram

```{r}
theme_set(theme_classic())

g = ggplot(mpg,aes(displ)) + scale_fill_brewer(palette = "Spectral")

g + geom_histogram(aes(fill=class),
                   binwidth = .1,
                   col = "black",
                   size = .1) +
  labs(title = "Histogram with Auto Binning" ,
       subtitle = "Engine Displacement across Vehicle Classes")

g + geom_histogram(aes(fill=class),
                   bins = 5,
                   col = "black",
                   size = .1) +
  labs(title = "Histogram with Fixed Bins",
       subtitle = "Engine Displacement across Vehicle Classes")
```

#### Histogram on a categorical variable

```{r}
theme_set(theme_classic())

g = ggplot(mpg,aes(manufacturer))

g + geom_bar(aes(fill=class),width=0.5) +
  theme(axis.text.x = element_text(angle=65,vjust=0.6)) +
  labs(title = "Histogram on Categorical Variable" ,
       subtitle = "Manufacture across Vehicle Classes")

```
#### Density plot

```{r}
theme_set(theme_classic())

g = ggplot(mpg,aes(cty))

g + geom_density(aes(fill=factor(cyl)),alpha=0.8) +
  labs(title = "Density plot" ,
       subtitle = "City Mileage Grouped by Number of cylinders",
       x = "City Mileage",
       fill = "# Cylinders")

```
#### Box plot

```{r}
theme_set(theme_classic())

g = ggplot(mpg,aes(class,cty))

g + geom_boxplot(varwidth = T, fill="Plum") +
  labs(title="Box plot",
       subtitle = "City Mileage grouped by Class of vehicle",
       caption = "Source:mpg",
       x = "Class of Vehicle",
       y = "City Mileage")

```

#### Dot + Box plot

```{r}
theme_set(theme_bw())

g = ggplot(mpg,aes(manufacturer,cty))

g + geom_boxplot() +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "red") +
  theme(axis.text.x = element_text(angle=65,vjust=.6)) +
  labs(title="Box plot + Dot plot",
       subtitle = "City Mileage vs Class: Each dot represents 1 row in source data",
       caption = "Source:mpg",
       x = "Class of Vehicle",
       y = "City Mileage")

```

#### Tufte Boxplot

```{r}
library(ggthemes)
theme_set(theme_tufte())

g = ggplot(mpg,aes(manufacturer,cty))

g + geom_tufteboxplot() +
  theme(axis.text.x = element_text(angle=65,vjust=.6)) +
  labs(title="Tufte Styled Boxplot",
       subtitle = "City Mileage grouped by Class of vehicle",
       caption = "Source:mpg",
       x = "Class of Vehicle",
       y = "City Mileage")

```













