---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Intro I

### 1. Understand

#### setup
```{r}
options(scipen = 999)
library(ggplot2)
data("midwest",package="ggplot2")
```

#### Init Ggplot
aes() function is used to specify the X and Y axes
```{r}
ggplot(midwest,aes(x=area,y=poptotal))
```

### 2. Make a Simple Scatterplot

#### Add Scaterplot
```{r}
ggplot(midwest,aes(x=area,y=poptotal)) + geom_point()
```
#### Add smoothing layer
the method "lm" is the short for linear model and it draws the line of best fit
```{r}
ggplot(midwest,aes(x=area,y=poptotal)) + geom_point() + geom_smooth(method="lm")
```

### 3. Adjusting X and Y axis limits

#### Method1: By deleting the points outside the range
It can be down with xlim() ylim()
When using xlim() and ylim(), the points outside the specificed range are deleted and will not be considered while drawing the line of best fit (the method "lm")
```{r}
g = ggplot(midwest,aes(x=area,y=poptotal)) + geom_point() + geom_smooth(method="lm")

g + xlim(c(0,0.1)) + ylim(c(0,1000000)) # It will be same without c( , )
```
#### Method2: Zooming In
It can be down with coord_cartesian()
```{r}
g = ggplot(midwest,aes(x=area,y=poptotal)) + geom_point() + geom_smooth(method="lm")

g1 = g + coord_cartesian(xlim=c(0,0.1),ylim=c(0,1000000))
plot(g1)
```
### 4. Change the Title and Axis Labels

Function labs() is for title, x, and y;

```{r}
g1 +labs(title = "Area Vs Population", subtitle = "From midwest dataset", y = "Population", x = "Area", caption = "Midwest Demographics")

```
Or ggtitle(), xlab(), ylab()
```{r}
g1 + ggtitle("Area Vs Population", subtitle = "From midwest dataset") + xlab("Area") + ylab("Population") 
```

### 5. Change the Color and Size of Points

#### Static 
```{r}
ggplot(midwest,aes(x=area,y=poptotal)) +
  geom_point(col="steelblue",size=3) + # set static color and size for points
  geom_smooth(method="lm",col="firebrick") + # change the color of line
  coord_cartesian(xlim = c(0,0.1), ylim=c(0,1000000)) +
  labs(title="Area Vs Population", subtitle = "From midwest dataset", y = "Population", x = "Area", caption = "Midwest Domographics")
```
#### Adaptive: Change the color to reflect categories in another column
```{r}
gg = ggplot(midwest,aes(x=area,y=poptotal)) +
  geom_point(aes(col=state),size=3) + # set color to vary based on state categories
  geom_smooth(method="lm",col="firebrick",size=2) + # change the color of line
  coord_cartesian(xlim = c(0,0.1), ylim=c(0,1000000)) +
  labs(title="Area Vs Population", subtitle = "From midwest dataset", y = "Population", x = "Area", caption = "Midwest Domographics")

plot(gg)
```
aes(col=state) means the color is based on state.Size, shape, stroke and fill(fill color) can also be used to discriminate groupings.

Legend is added automatically. It can be removed by setting "legend.position" to "None" from within a theme() function:
```{r}
gg+theme(legend.position = "None")
```
Change the color palette entirely:
```{r}
gg+scale_colour_brewer(palette="Set1") # Change the color palette
```
More color palettes can be found in the package "RColorBrewer"
```{r color, echo=FALSE}
library(RColorBrewer)
head(brewer.pal.info,10) # show 10 palettes
```

### 6. Change the X and Y Axis Texts and Ticks Location

#### Step1: Set the breaks:
```{r}
gg + scale_x_continuous(breaks = seq(0,0.1,0.01))
```
#### Step2: Change the labels
```{r}
gg + scale_x_continuous(breaks = seq(0,0.1,0.01),labels = letters[1:11])
```
if we want to reverse X Axis Scale
```{r}
gg + scale_x_reverse()
```
#### Customized Texts for Axis Labels
```{r}
gg + 
  scale_x_continuous(breaks=seq(0,0.1,0.01),labels=sprintf("%1.2f%%",seq(0,0.1,0.01))) +
  scale_y_continuous(breaks=seq(1,1000000,200000),labels=function(x)(paste0(x/1000,'K')))
```
#### Customized the Entire Theme in One Shot using Pre-BUilt Themes
if we want to reverse X Axis Scale
```{r}
gg = gg + scale_x_continuous(breaks = seq(0,0.1,0.01))
```
method 1: Using theme_set()
```{r}
theme_set(theme_classic())
gg
```
method 2: Adding theme Layer itself
```{r}
gg + theme_bw() + labs(subtitle = "BW Theme")
gg + theme_classic() + labs(subtitle = "Classic Theme")
gg
```

## Intro II

### 1. Adding Plot and Axis Titles

