---
title: "Ggplot_part2"
author: "Zhang Xiaoqiong"
date: "1/20/2020"
output:
  pdf_document: default
  html_document: default
---
Note: It also requires the "ggrepel" and "gridExtra" packages. please download them first to run the code
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part2: How to customize ggplot2


```{r setups}
options(scipen=999)
library(ggplot2)
data('midwest',package = 'ggplot2')
theme_set(theme_bw())
```


```{r add plot components}
gg = ggplot(midwest,aes(x=area,y=poptotal)) +
  geom_point(aes(col=state,size=popdensity)) +
  geom_smooth(method = 'loess',se=F) + xlim(c(0,0.1)) + ylim(c(0,500000)) +
  labs(title='Area Vs Population',y = 'Population',x = 'Area',caption = 'Sourse: midwest')
plot(gg)
```

### 1. Adding Plot and Axis Titles

```{r }
gg + theme(plot.title = element_text(size=20,
                                     face="bold", # set fond face
                                     family="American Typewriter", # a new font
                                     color="tomato",
                                     hjust=0.5,
                                     lineheight=1.2),
           plot.subtitle = element_text(size=15,
                                        family="American Typewriter",
                                        face="bold",
                                        hjust=0.5), # controls horizontal spacing
           plot.caption = element_text(size=15),
           axis.title.x = element_text(vjust=1,   # controls vertical spacing
                                     size=15),
           axis.title.y = element_text(size=15),
           axis.text.x = element_text(size=10,
                                     angle=30,
                                     vjust=.5),
           axis.text.y = element_text(size = 10))

```

### 2. Modifying Legend

Method 1: use labs()
```{r }
gg + labs(color="State",size="Density") # modify legend title

```
Method 2: use guides()
```{r }
gg + guides(color=guide_legend("State"),size=guide_legend("Density"))

```
Method 3: use scale_aesthetic_vartype(). It can turn off legend for one particular aesthetic, leaving the rest in place
```{r }
gg + scale_color_discrete(name="State") + scale_size_continuous(name="Density",guide = FALSE) # turn off legend for size
```

#### How to change legend labels and point colors for categories
```{r}
gg + scale_color_manual(name="State",
                        labels=c("Illinois",
                                 "Indiana",
                                 "Michigan",
                                 "Ohio",
                                 "Wisconsin"),
                        values = c("IL" = "blue",
                                   "IN" = "red",
                                   "MI" = "green",
                                   "OH" = "brown",
                                   "WI" = "organfe"))
```
#### Change the Order of Legend
```{r}
gg + guides(colour = guide_legend(order=1),
            size = guide_legend(order=2))
```

#### How to style the legend title, text and key
```{r}
gg + theme(legend.title = element_text(size=12,color="firebrick"),
           legend.text = element_text(size=10),
           legend.key = element_rect(fill='springgreen')) +
  guides(colour = guide_legend(override.aes = list(size=2,stroke=1.5)))
```

#### Remove the legend and change legend positions
```{r}
# No legend
gg + theme(legend.position = "None") + labs(subtitle = "No legend")

# Legend to the left
gg + theme(legend.position = "left") + labs(subtitle = "Legend on the Left")

# Legend at the bottom and horizontal
gg + theme(legend.position = "bottom", legend.box = "Horizontal") + labs(subtitle = "Legend at Bottom")

# Legend at bottom-right, inside the plot
gg + theme(legend.title = element_text(size=12,color="salmon",face="bold"),
           legend.justification = c(1,0),
           legend.position = c(0.95,0.05),
           legend.background = element_blank()) +
  labs(subtitle = "Legend: Bottom-Right Inside the Plot")

# Legend at top-left, inside the plot
gg + theme(legend.title = element_text(size=12,color="salmon",face="bold"),
           legend.justification = c(1,0),
           legend.position = c(0.05,0.95),
           legend.background = element_blank(),
           legend.key = element_blank()) +
  labs(subtitle = "Legend: Top-Left Inside the Plot")
```

### 3. Adding Text, Label and Annotation

#### Add text, label around the point
```{r }

# filter required rows
midwest_sub = midwest[midwest$poptotal > 300000,]
midwest_sub$large_county = ifelse(midwest_sub$poptotal > 300000,midwest_sub$county,"")

# plot text and label
gg + geom_text(aes(label=large_county),size=2,data=midwest_sub)+labs(subtitle="With ggplot2::geom_text") + theme(legend.position = "None") # text
gg + geom_label(aes(label=large_county),size=2,data=midwest_sub,alpha=0.25) + labs(subtitle="With ggplot::geom_label") + theme(legend.position = "None") # label

# plt text and label that repels eachother (using ggrepel pkg)
library(ggrepel)

gg + geom_text_repel(aes(label=large_county),size=2,data = midwest_sub) + labs(subtitle = "With ggrepel::geom_text_repel") + theme(legend.position = "None") # text
gg + geom_label_repel(aes(label=large_county),size=2,data = midwest_sub) + labs(subtitle = "With ggrepel::geom_label_repel") + theme(legend.position = "None") # label
```

#### Add Annotations Anywhere inside plot
```{r }
#define and add annotation
library(grid)
my_text = "This text is at x=0.7 and y=0.8!"
my_grob = grid.text(my_text,x=0.7,y=0.8,gp=gpar(col="firebrick",fontsize=14,fontface="bold"))
gg + annotation_custom(my_grob)
```




### 4. Flipping and Reversing X and Y axis

#### Reverse X and Y
```{r }
gg + labs(subtitle = "X and Y axis Flipped") + theme(legend.position = "None") + coord_flip()
```
#### Reverse the scale of an axis
```{R}
gg + labs(subtitle = "X and Y axis Reversed") + theme(legend.position = "None") + scale_x_reverse() + scale_y_reverse()
```
### 5. Faceting: Draw multiple plots within one figure
```{R}
data(mpg,package="ggplot2")

g = ggplot(mpg,aes(x=displ,y=hwy)) +
  geom_point() +
  labs(title="hwy vs disp",caption="Source:mpg") +
  geom_smooth(method="lm",se=FALSE) +
  theme_bw()
plot(g)
```

#### Face Wrap

facet_wrap is used to break down a large plot into multiple small plots for individual categories
```{R}
g + facet_wrap(~ class,nrow=3) + labs(subtitle="Ggplot2 - Faceting - Multiple plots in one figure") # Shared scales

g + facet_wrap(~ class,scales="free") + labs(title = "hwy vs displ",caption = "Source: mpg", subtitle = "Ggplot2 - Faceting - Multiple plots in one figure with free scales") # Scales free
```

#### Face Grid

facet_grid would get rid of unsignificant space.
```{R}
g1 = g + labs(subtitle="Ggplot2 - Faceting - Multiple plots in one figure - For Manufacturer") + facet_grid(manufacturer ~ class)
plot(g1)
```

```{R}
g2 = g + labs(subtitle="Ggplot2 - Faceting - Multiple plots in one figure - For Cylinder") + facet_grid(cyl ~ class)
plot(g2)
```

Layout both these charts in the sample panel

```{R}
library(gridExtra)
gridExtra::grid.arrange(g1,g2,ncol=2)

```
### 6. Modifying Plot Background, Major and Minor Axis

#### Change Plot background
```{R}
# Base plot
g = ggplot(mpg,aes(x=displ,y=hwy)) +
  geom_point() +
  geom_smooth(method="lm",se=FALSE) +
  theme_bw()

# Change plot background elements
g + theme(panel.background = element_rect(fill = 'khaki'),
          panel.grid.major = element_line(colour = "burlywood",size=1.5),
          panel.grid.minor = element_line(colour = "tomato",size=.25,linetype = "dashed"),
          panel.border = element_blank(),
          axis.line.x = element_line(colour="darkorange",
                                     size=1.5,
                                     lineend = "butt"),
          axis.line.y = element_line(colour="darkorange",
                                     size=1.5)) +
  labs(title="Modifies Background",
       subtitle="How to Change Major and Minor grid, Axis Lines, No Border")

# Change plot margins
g + theme(plot.background = element_rect(fill="salmon"),
          plot.margin = unit(c(2,2,1,1),"cm")) +
  labs(title="Modified Background",subtitle = "How to Change Plot Margin")
```
#### Remove Major and Minor Grid, Change Border, Axis Title, Text and Ticks
```{R}
g + theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
  labs(title="Modified Background", subtitle="How to remove major and minor axis grid, border, axis title, text and ticks")

```

#### Add image as background
```{R}
library(grid)
library(png)
img = png::readPNG("/Users/xiaoqiongzhang/Downloads/Rlogo.png") #Source: https://www.r-projects.org/  The logo is on the top left side.
g_pic = rasterGrob(img,interpolate = TRUE)

g + theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.title = element_text(size=rel(1.5),face="bold"),
          axis.ticks = element_blank()) +
  annotation_custom(g_pic,xmin=5,xmax=7,ymin=30,ymax=45)
```





